---
title: 'Homework #7'
author: "Collin Rafferty"
date: "11/10/2021"
output:
  word_document: default
  pdf_document: default
---
```{r}
load("/cloud/project/Household_Pulse_data.RData")

Household_Pulse_data$vaxx <- (Household_Pulse_data$RECVDVACC == "yes got vaxx")
is.na(Household_Pulse_data$vaxx) <- which(Household_Pulse_data$RECVDVACC == "NA") 

# Like last week, I decided to set the NAs equal to zero. 

require(gtsummary)
tb1 <- data.frame(Col = Household_Pulse_data$vaxx,
 Row = Household_Pulse_data$EEDUC)
tbl_cross(tb1, row=Row, col=Col, percent="row")

tb2 <- data.frame(Col = Household_Pulse_data$vaxx,
 Row = Household_Pulse_data$RRACE)
tbl_cross(tb2, row=Row, col=Col, percent="row")

tb3 <- data.frame(Col = Household_Pulse_data$vaxx,
 Row = Household_Pulse_data$REGION)
tbl_cross(tb3, row=Row, col=Col, percent="row")

pick_use1 <- (Household_Pulse_data$REGION == "South" & Household_Pulse_data$GENID_DESCRIBE== "male" & Household_Pulse_data$RRACE== "White")
dat_use1 <- subset(Household_Pulse_data, pick_use1)
```

I thought it would be interesting to look at a subset of only white males living in the south because, in lab 6, I found them to have the lowest probability of being vaccinated out of the subset I was looking at.

```{r}
d_marstat <- data.frame(model.matrix(~ dat_use1$MS))
d_pubhlth<- data.frame(model.matrix(~ dat_use1$PUBHLTH))
d_x <- data.frame(model.matrix(~ dat_use1$SEXUAL_ORIENTATION))
d_income <- data.frame(model.matrix(~ dat_use1$INCOME))
d_educ <- data.frame(model.matrix(~ dat_use1$EEDUC))
d_vaxx <- data.frame(model.matrix(~ dat_use1$vaxx))

dat_for_analysis_sub <- data.frame(
  d_vaxx[,2],
  d_educ[!is.na(dat_use1$vaxx),2:7],
  d_marstat[!is.na(dat_use1$vaxx),2:6],
  d_income[!is.na(dat_use1$vaxx),2:9],
  d_pubhlth[!is.na(dat_use1$vaxx),2:3],
  d_x[!is.na(dat_use1$vaxx),2:6])

names(dat_for_analysis_sub) <- sub("dat_use1.","",names(dat_for_analysis_sub))
names(dat_for_analysis_sub)[1] <- "vaxx"

summary(d_vaxx)
summary(d_educ)
```

```{r}
require("standardize")
set.seed(12345)
NN <- length(dat_for_analysis_sub$vaxx)
restrict_1 <- (runif(NN) < 0.3) 
summary(restrict_1)
dat_train <- subset(dat_for_analysis_sub, restrict_1)
dat_test <- subset(dat_for_analysis_sub, !restrict_1)

summary(dat_test$INCOMEHH.income..200k..)

sobj <- standardize(vaxx ~  EEDUCsome.hs + EEDUCHS.diploma + EEDUCsome.coll + EEDUCassoc.deg + EEDUCbach.deg + EEDUCadv.deg + MSmarried + MSwidowed + MSdivorced + MSseparated +PUBHLTHno.public.health.ins+ MSnever+INCOMEHH.income.less.than..25k+INCOMEHH.income..25k....34.9k+INCOMEHH.income..35k...49.9+ INCOMEHH.income..50k...74.9+INCOMEHH.income..75...99.9+INCOMEHH.income..100k...149+INCOMEHH.income..150...199+ SEXUAL_ORIENTATIONbisexual+ SEXUAL_ORIENTATIONdont.know+SEXUAL_ORIENTATIONgay.or.lesbian+SEXUAL_ORIENTATIONsomething.else+SEXUAL_ORIENTATIONstraight, dat_train, family = binomial)

summary(sobj$data)
s_dat_test <- predict(sobj, dat_test)
```
After running a summary on incomes over 200K, I found its Min=Max=0, so I dropped it from the model. I didn't want it to cause an error in the model. 
```{r}
require(stargazer)
model_lpm1 <- lm(sobj$formula, data = sobj$data)
stargazer::stargazer(model_lpm1, type="text", title = "Model_lmp1")
pred_vals_lpm <- predict(model_lpm1, s_dat_test)
pred_model_lpm1 <- (pred_vals_lpm > mean(pred_vals_lpm))
table(pred = pred_model_lpm1, true = dat_test$vaxx)
# logit 
model_logit1 <- glm(sobj$formula, family = binomial, data = sobj$data)
stargazer::stargazer(model_logit1, type="text", title = "Model_logit1")
pred_vals <- predict(model_logit1, s_dat_test, type = "response")
pred_model_logit1 <- (pred_vals > 0.5)
table(pred = pred_model_logit1, true = dat_test$vaxx)
pred_model_logit2 <- (pred_vals > 0.3)
table(pred = pred_model_logit2, true = dat_test$vaxx)
pred_model_logit3 <- (pred_vals > 0.8)
table(pred = pred_model_logit3, true = dat_test$vaxx)

cm1 <- as.table(matrix(c(6, 5, 504, 4278), nrow = 2, byrow = TRUE))
fourfoldplot(cm1, color = c("#CC6666", "#99CC99"),
conf.level = 0, margin = 1, main = "Confusion Matrix for Model_logit1 ")

cm2 <- as.table(matrix(c(0, 1, 510, 4283), nrow = 2, byrow = TRUE))
fourfoldplot(cm2, color = c("#CC6666", "#99CC99"),
conf.level = 0, margin = 1, main = "Confusion Matrix for Model_logit2 ")

cm3 <- as.table(matrix(c(210, 597, 300, 3686), nrow = 2, byrow = TRUE))
fourfoldplot(cm3, color = c("#CC6666", "#99CC99"),
conf.level = 0, margin = 1, main = "Confusion Matrix for Model_logit3 ")

```
The tables and graphs show as the model's cutoff value is reduced, the model's predictive accuracy increases to a point. For example, when the cutoff value was set to pred_vals > .8, the model made 4283 true positive predictions, but when the cutoff value was set to pred_vals > .5, the model only predicted 3686 true positives. Both false negative and false positive also increase substantially from model one to model three. However, there is a certain point when the cutoff value is too low, and the model's predictive accuracy declines. For example, looking at model two, where the cutoff was set to pred_vals >.3, it predicted 4278 true positives. This is five fewer true positives than model one, in which the cutoff value was set to pred_vals >.5. 