---
title: 'Homework #2'
author: "Collin Rafferty"
date: "9/24/2021"
output: html_document
---

1. My group members were Nicholas Alonso, Adelia Fida, and Serigne Cisse. 

2. After looking through the PUMS data, I decided to compare the degree fields of computer and information sciences to the fine arts. 
I thought these two-degree fields would be the stark opposite of one another, so it would reveal some interesting facts about the data. 
I began by comparing some basic stats between the two-degree fields.  

```{r}
mean(as.numeric(acs2017_ny$HHINCOME[(acs2017_ny$DEGFIELD=="Computer and Information Sciences" &acs2017_ny$female==0)]), na.rm = TRUE)
mean(as.numeric(acs2017_ny$HHINCOME[(acs2017_ny$DEGFIELD=="Fine Arts" &acs2017_ny$female==0)]), na.rm = TRUE)
table(acs2017_ny$Hispanic, acs2017_ny$DEGFIELD=="Computer and Information Sciences")
table(acs2017_ny$Hispanic, acs2017_ny$DEGFIELD=="Fine Arts")
```

Going over these basic stats raised an interesting question: how does the average total household 
income of a person of Hispanic origin compare to the larger population surveyed? In a study from the 
Economic Policy Institute, Mora and Davila found college-educated Hispanic males made, on average, 14.9% 
less per hour than college-educated white males. Due to the finite time I had to analyze this data, I decided to look 
at how the average total household income of Hispanic males with degrees in computer and information sciences compares to the overall male 
population with degrees in computer and information sciences.

```{r}
restrict1 <- (acs2017_ny$Hispanic == 0)| (acs2017_ny$female==0) | (acs2017_ny$DEGFIELD=="Computer and Information Scienes")
data_new1 <- subset(acs2017_ny,restrict1)
restrict2 <- (acs2017_ny$Hispanic == 1) | (acs2017_ny$female==0) | (acs2017_ny$DEGFIELD=="Computer and Information Scienes")
data_new2 <- subset(acs2017_ny,restrict2)
mean(as.numeric(data_new1$HHINCOME), na.rm = TRUE)
mean(as.numeric(data_new2$HHINCOME), na.rm = TRUE)
error1 <- qt(0.975,df=length(data_new1) -1)*sd(as.numeric(data_new1$HHINCOME), na.rm = TRUE)/sqrt(length (data_new1))
error2 <- qt(0.975,df=length(data_new2) -1)*sd(as.numeric(data_new2$HHINCOME), na.rm = TRUE)/sqrt(length (data_new2))
left1 <- mean(as.numeric(data_new1$HHINCOME), na.rm = TRUE)-error1
right1 <- mean(as.numeric(data_new1$HHINCOME), na.rm = TRUE)+error1
left2 <- mean(as.numeric(data_new2$HHINCOME), na.rm = TRUE)-error2
right2 <- mean(as.numeric(data_new2$HHINCOME), na.rm = TRUE)+error2
```
I am 95% confident that the true mean for non-Hispanic males with degrees in computer and information sciences is within 
the interval between $32,560 and $40,610. Likewise, I am 95% confident that the true mean for males with degrees 
in computer and information sciences is within the interval between $32,120 and $40,150.  

```{r}
table(data_new1$educ_advdeg, data_new1$AGE < 30)
table(data_new2$educ_advdeg, data_new2$AGE < 30)
prop.table(table(data_new1$educ_advdeg, data_new1$AGE < 30))
prop.table(table(data_new2$educ_advdeg, data_new2$AGE < 30))
```

In the crosstabs, I examined the percentage of Hispanic males under the age of 30 with advanced degrees in 
computer and information sciences compared to non-Hispanic males of the same age and educational attainment. The data showed 
that non-Hispanic males were .3% more likely to be under 30 with an advanced degrees. This is not a 
large percentage, but I believe it shows a pattern. If non-Hispanic males are more likely to get advanced degrees 
in this field, this will translate to them having greater salaries. These differences will cause more inequality in pay and opportunities 
between Hispanic and non-Hispanic males in this field.

Marginal Probabilities using Bayes’ Theorem

Tad= Holds an advanced degree ; Fad= Does not Holds an advanced degree ; Y30 = younger than 30 ;O30 = older than 30 
P(Tad|Y30)= P(Y30|Tad)*P(Tad)/ P(Y30) = (1999/23385)*(66429/196585)/ (23385/196585)=0.2428263
 P(Fad|O30)= P(O30|Fad)*P(Fad)/ P(O30) = (21386/23385)*(130156/196585)/ (173200/196585)=0.6872401
0.2428263+0.6872401 ≈ 1

The crosstabs I constructed was mutually exclusive and exhaustive. First, if someone had an advanced degree, they 
cannot be grouped into thecolumn of people who don't have advanced degrees. That just wouldn't make sense. The same goes for
age. Second, the crosstab is exhaustive because one event always occurs: the person has or does not have the advanced degree; the person 
is under the age of 30 or over the age of 30. 

A potential confounding variable that could be affecting the results I received is citizenship status. For example, if a Hispanic 
person grew up in a different country speaking Spanish and then came to the U.S. for university or employment, 
they may not have as good a grasp of English as someone born and raised in the U.S. does. This could explain some of 
the salaries and advance degree attainment differences. 

If I could examine additional evidence beyond this dataset, I would like to look at a breakdown of salaries difference 
between Hispanic vs. non-Hispanic males by different companies. I would posit that companies founded relatively recently, 
last 30 years, like Google, Amazon, and others, the salary disparity would be far less pronounced than 
in older tech companies - IBM and others. I think this would be because of the ingrained biases these older companies have. If 
I were right about this disparity, it would make a persuasive case on why things need to change. 
