---
title: "Homework # 3"
author: "Collin Rafferty"
date: "9/30/2021"
output: html_document
---
Based on the insights I gleaned from last week's homework, I decided to look at how the Hispanic population of New York is spread out 
across the boroughs. I decided to look at Hispanic people who did not have any form of American citizenship and whose 
household makes less than $30,000 annually. Also, whose family sizes are larger than 5 members. The inhumane treatment of delivery 
workers has been all over the news over the past few weeks, so I believe this analysis will give some helpful insights into the 
geographic distribution of some of these workers. 

```{r}
# k-nn classification
dat_NYC <- subset(acs2017_ny, (acs2017_ny$in_NYC == 1)&(acs2017_ny$Hispanic=1) & (acs2017_ny$CITIZEN=5))
attach(dat_NYC)
borough_f <- factor((in_Bronx + 2*in_Manhattan + 3*in_StatenI + 4*in_Brooklyn + 5*in_Queens), levels=c(1,2,3,4,5),labels = c("Bronx","Manhattan","Staten Island","Brooklyn","Queens"))
norm_varb <- function(X_in) {(X_in - min(X_in, na.rm = TRUE))/
( max(X_in, na.rm = TRUE) - min(X_in, na.rm = TRUE) )}
is.na(HHINCOME) <- which(HHINCOME <= 30000)
is.na(FAMSIZE) <- which(FAMSIZE >=5)
norm_HHINCOME <- norm_varb(HHINCOME)
norm_FAMSIZE <- norm_varb(FAMSIZE)
```
```{r}
# Creation of the dataframe
data_use_prelim <- data.frame(norm_HHINCOME,norm_FAMSIZE)
good_obs_data_use <- complete.cases(data_use_prelim,borough_f)
dat_use <- subset(data_use_prelim,good_obs_data_use)
y_use <- subset(borough_f,good_obs_data_use)

# 80/20 split of data into training set and validation set
set.seed(12345)
NN_obs <- sum(good_obs_data_use == 1)
select1 <- (runif(NN_obs) < 0.8)
train_data <- subset(dat_use,select1)
test_data <- subset(dat_use,(!select1))
cl_data <- y_use[select1]
true_data <- y_use[!select1]

# Running the k-nn algorithm to be compared to the subset's means
summary(cl_data)
prop.table(summary(cl_data))
summary(train_data)
require(class)
for (indx in seq(1, 9, by= 2)) {
pred_borough <- knn(train_data, test_data, cl_data, k = indx, l = 0, prob = FALSE, use.all = TRUE)
num_correct_labels <- sum(pred_borough == true_data)
correct_rate <- num_correct_labels/length(true_data)
print(c(indx,correct_rate))}
```
```{r}
# Simple linear regression
cl_data_n <- as.numeric(cl_data)

model_ols1 <- lm(cl_data_n ~ train_data$norm_HHINCOME + train_data$norm_FAMSIZE)

y_hat <- fitted.values(model_ols1)

mean(y_hat[cl_data_n == 1])
mean(y_hat[cl_data_n == 2])
mean(y_hat[cl_data_n == 3])
mean(y_hat[cl_data_n == 4])
mean(y_hat[cl_data_n == 5])
```
It is interesting that close to 67% of the population we are looking at live in either Brooklyn(34%)
or Queens(33%). While only about 16% live in Manhattan, and the other roughly 17% live in the Bronx(11%) and Staten Island(6%). 
Intuitively this makes sense when you think about it. 

For the 2nd Iteration of the knn-classification, I restricted the age to 30 or younger because I think most delivery workers are in this age range. 

```{r}
is.na(AGE) <- which(AGE<= 30)
norm_AGE <- norm_varb(AGE)

# Creation of the dataframe
data_use_prelim <- data.frame(norm_HHINCOME,norm_FAMSIZE,norm_AGE)
good_obs_data_use <- complete.cases(data_use_prelim,borough_f)
dat_use1 <- subset(data_use_prelim,good_obs_data_use)
x_use <- subset(borough_f,good_obs_data_use)

# 80/20 split of data into training set and validation set
set.seed(12345)
NN_obs <- sum(good_obs_data_use == 1)
select2 <- (runif(NN_obs) < 0.8)
train_data1 <- subset(dat_use1,select2)
test_data1 <- subset(dat_use1,(!select2))
cl_data1 <- x_use[select2]
true_data1 <- x_use[!select2]

# Running the k-nn algorithm to be compared to the subset's means
summary(cl_data1)
prop.table(summary(cl_data1))
summary(train_data1)
require(class)
for (indx in seq(1, 9, by= 2)) 
pred_borough <- knn(train_data, test_data1, cl_data1, k = indx, l = 0, prob = FALSE, use.all = TRUE)
num_correct_labels1 <- sum(pred_borough == true_data1)
correct_rate1 <- num_correct_labels1/length(true_data1)
print(c(indx,correct_rate))
```
It is interesting how the second iteration flipped the percentage of Queens(now 34%) and Brooklyn(now 34%). 
Manhattan's percentage dropped to 15%, and the other two boroughs stayed about the same. 

For the 3rd and final iteration of the knn-classification, I restricted rent to $1000 or less to see how this affected the geographic distribution. 

```{r}
is.na(RENT) <- which(RENT<= 1000)
norm_RENT <- norm_varb(RENT)

# Creation of the dataframe
data_use_prelim <- data.frame(norm_HHINCOME,norm_FAMSIZE,norm_RENT)
good_obs_data_use <- complete.cases(data_use_prelim,borough_f)
dat_use2 <- subset(data_use_prelim,good_obs_data_use)
z_use <- subset(borough_f,good_obs_data_use)

# 80/20 split of data into training set and validation set
set.seed(12345)
NN_obs <- sum(good_obs_data_use == 1)
select3 <- (runif(NN_obs) < 0.8)
train_data2 <- subset(dat_use1,select3)
test_data2 <- subset(dat_use1,(!select3))
cl_data2 <- z_use[select3]
true_data2 <- z_use[!select2]

# Running the k-nn algorithm to be compared to the subset's means
summary(cl_data2)
prop.table(summary(cl_data2))
summary(train_data2)
require(class)
for (indx in seq(1, 9, by= 2)) 
pred_borough <- knn(train_data, test_data1, cl_data1, k = indx, l = 0, prob = FALSE, use.all = TRUE)
num_correct_labels2 <- sum(pred_borough == true_data2)
correct_rate2 <- num_correct_labels2/length(true_data2)
print(c(indx,correct_rate))
```
The 3rd iteration had a pretty significant impact on the geographic distribution of the subset population. First, Brooklyn went up to 36%, 
while Queens went down to 30%. Manhattan increased to 20%. The Bronx stayed at about the same level, but Staten island fell all the way to .02%. 

I think the tradeoff between the better classification of the training data and the worsening of the model's ability to classify 
the testing data occurs around a ratio of 90/10. If you devote more than 90% of the data set to train the model, the model 
will become biased towards the particular dataset. This will make it hard for the model to become generalized when it is looking at unseen data. 

I think the 3 iterations, which were ran, provided a good amount of helpful insight into how one of the most unprivileged groups in New York City is geographically distributed. 

